---
title: "README"
author: "Edson Nilton de Moura SIlva Júnior"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    number_sections: no
    fig_caption: yes
    theme: flatly
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Dealing with outliers in R

Outliers can be defined by valeus or ammostraal units in a vector or sample which are highlightly distinct from the others. They can be a trouble to many statistics analysis, thus outliers affect them, increasing the chances to occour Type 1 Error (reject null hypothesis when it is true). Based this, detecting and knowing outliers in a vector, such a variable, can be high usefull to data science.

# Required packages

To initialize the analysis, it is required the `tidyverse` package. 

```{r}
library(tidyverse)
```

# Dataset

To do the analyses, it will use the `Iris dataset`, a dataset containg data about three species from *Iris* genera: *I. setosa*, *I. versicolor*, and *I. virginica*. That dataset have data about their sepal length and width, and petal length and width.

```{r}
data("iris")

iris
```

# Load functions

Three function was build to deal about outliers: `outlier_detect()`, to detect the the presence of upper and lower outliers; `outlier_plot()`, to plot the vector in a ggplot scatter plot, vizualizing their distribuition; and `outlier_percent()`, to calculate the percentage of upper and lower outlier in the vector. The fuctions were build and saved in .R files. check them in repository files.

To load the functions, it was used the `source()` function to that three .R files, containg the functions.

```{r}
source("outlier detect.r")

source("outlier plot.r")

source("outlier percente.r")
```

# Detectig

```{r}
for(i in iris[1:4] %>% names()){
  
  str_glue("# detecting outlier to {i}") %>% message()
  
  outlier_detect(iris[i] %>% pull()) 
  
  message(" ")
  
}
```

# Ploting

```{r}
for(i in iris[1:4] %>% names()){
  
  str_glue("# plotting outlier to {i}") %>% message()
  
  outlier_plot(iris[i] %>% pull()) %>% print()
  
  message(" ")
  
}
```

# Percenatge

```{r}
for(i in iris[1:4] %>% names()){
  
  str_glue("# Percentage outlier to {i}") %>% message()
  
  outlier_percent(iris[i] %>% pull())
  
  message(" ")
  
}
```

